Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2016-07-20 manuscript/manuscript_pve.Rmd
Amass clarification
final edits, submitted this version to new phyt

\
In the absence of a good understanding of the effects of sink limitation on physiological processes, most attempts to model sink limitation of growth to date have done so by abandoning the principle of mass balance [e.g. @fatichi2013reconciling; @fatichi2014moving]. This approach is unsatisfactory because mechanistic models should obey the laws of physics. Growth models must take into account several simultaneous processes affecting the C balance at any given time [@fourcaud2008plant], thus incomplete modelling approaches often lead to unaccounted-for C. Our goal in this paper was to investigate how the C balance of plants is modified by sink limitation, by experimentally imposing a sink limitation to growth.

@@

\
Forty-nine *Eucalyptus tereticornis* seedlings, 20 weeks old and approximately 40 cm tall in tube stock, were chosen from a single local Cumberland plain (western Sydney) cohort. Six additional seedlings were harvested before planting to measure initial leaf area and dry mass of leaves, stems and roots. Six container volumes were used ranging from 5 L to 35 L, with a 22.5 cm diameter, and lengths ranging from 15 to 100 cm. Containers were constructed of PVC pipe and were filled with local top soil. Soil in each container was packed to achieve a target soil bulk density that matched local soil conditions of 1.7 g cm^-3^. An Imidacloprid (BAYER CropScience) insecticide tablet was planted 5 cm below the roots of each seedling. Containers were planted flush with the soil surface inside metal sleeves, designed to minimize excess air space between the container and outside soil while also allowing for container removal. This allowed for soil temperatures in containers to reflect conditions of naturally planted (free) seedlings. Each experimental block (n=7) contained a complete replicate set of six container volumes as well as one free seedling, with 1 m^2^ spacing. For each free seedling, used as the control, a 1 m^2^ subplot was excavated to the hard layer and replaced with the same soil used in each container. A border of root exclusion material was buried 0.25 m deep and extended 0.25 m above the ground surface around each subplot to exclude local vegetation, which was further kept out by periodic weeding.

@@

##Growth and morphology metrics
Seedlings were planted in summer (January 21^st^ 2013) and stem height, diameter at 15 cm and leaf count were measured weekly thereafter. After four months, a full biomass harvest was completed and the experiment ended (May 21^st^ 2013). Dry mass of leaves, stems, roots and total leaf area (LI-3100C Area Meter; LI-COR, Lincoln, NE, USA) were measured for each seedling. Mean individual leaf area for each harvested seedling was calculated by dividing total measured leaf area by total leaf count of only fully expanded leaves. Mean individual leaf area was then used to interpolate total seedling leaf area through time with weekly leaf counts. Root mass was collected by removing the roots system and passing soil from each container through a 1 mm sieve, washing, separating into fine and coarse roots (<2 mm and >2 mm diameter, respectively) and then drying to a constant mass. Roots of seedlings in containers were not considered pot bound, as matted roots along the soil-container interface were not observed. Roots from the free seedlings were collected by excavating each 1 m^2^ subplot to the hard layer and keeping only roots within the subplot. For each seedling, a sub-sample of washed fine roots was analyzed for root length using WhinoRhizo software (Regent Instruments Inc., Quebec, QC, Canada). Specific root length (SRL) is reported as the root length divided by the dry mass of each sub-sample (m g^-1^). Fine root length density (FRLD) for seedlings in containers is reported as the total fine root length divided by the volume of each container (m dm^-3^).

@@

##Growth and morphology
Plant height, diameter and leaf area diverged between container volumes soon after start of the experiment (Figure `r allomfig`a-c). First to diverge was seedling leaf area, which significantly differed among soil volumes (*P* < 0.029) during the 5^th^ week of the experiment. Significant differences were then observed in both height (8^th^ week) and then diameter (9^th^ week) significantly among soil volumes (*P* < 0.045 & 0.035, respectively). The large reductions in height gain and total leaf area in smaller compared to larger containers continued throughout the experiment. In this field study, colder temperatures and reductions in total PPFD per day (Figure `r airvarsfig`) likely slowed the growth of seedlings in the final weeks of the experiment. Seedlings maintained diameter growth throughout the experiment, although growth was marginal in smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (*P* < 0.001). Increases in both final stem diameter and cumulative leaf area (both *P* < 0.001) were found with increasing soil volume. We did not find a continuous effect of container size; rather, plant size was similar among smaller containers, and differences were driven mainly by the largest container and the free seedling treatments.

\
Total seedling biomass at harvest was significantly different across container volumes (*P* < 0.001) and between container treatments and free seedlings (*P* < 0.001, Table `r pvetab1`). On average, harvested biomass of free seedlings was 84% higher than that of seedlings in containers. Plant biomass was positively correlated with total leaf area across all treatments (*R*^2^ = 0.97, *P* < 0.001). Differences in biomass partitioning to leaves, stems and roots were not different across soil volumes after variation in seedling biomass across treatments was factored in the analysis (Figure `r allocfig`a,b). The final harvest root:shoot biomass ratio and  was conserved across all treatments, with a slightly higher shoot than root mass ($\overline{x}$ = 0.904, 95% CI = [0.846,1.119]). The ratio of leaf to fine root mass was also not different between treatments (Figure `r allocfig`c).

\
Overall, SRL was higher in seedlings in containers compared to free seedlings but only in some of the container size treatments (Table `r pvetab2`, *P* = 0.009). Fine root length density was significantly higher in the two smallest container sizes and was the lowest in the largest container size (Table `r pvetab2`, *P* < 0.001). Over the duration of the experiment SLA~f~ was higher in free seedlings, but was not different across container sizes (Table `r pvetab1`, *P* < 0.001) and this pattern was evident beginning in the first gas exchange measurement campaign (*P* < 0.001).

##Leaf and root chemistry
Leaf N~f~ was significantly higher in free seedlings and the largest container volume compared to the smaller container volumes at the onset of gas exchange measurements (6th week, *P* < 0.001). Throughout the remainder of the experiment the smallest container volume had significantly lower leaf N~f~ than other soil volumes, while free seedlings maintained the highest leaf N~f~ (Table `r pvetab1`, *P* < 0.001). Leaf starch content in the smallest container was ca. double  that of free seedlings (*P* = 0.039), while leaf soluble sugars did not differ across treatments throughout the experiment (Table `r pvetab1`). Differences in leaf starch between the free seedling and the smallest container were evident during the first gas exchange campaign (*P* = 0.001). Root N was higher in free seedlings compared to seedlings in containers but only for some of the container size treatments (Table `r pvetab2`).

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ (Figure `r asatfig`) and *A*~max~ (Table `r pvetab3`) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf starch and leaf N~f~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis was highly correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~max~ to either leaf N~f~ or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~max~ was higher when leaf N~f~ was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~max~  was also lower when leaf starch was high as higher leaf N~f~ often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, *A*~max~ was positively correlated with final harvest biomass across all seedlings (*P* < 0.001).

\
Both *J*~max~ and *V*~cmax~ were significantly higher in free seedlings (30 % and 26 %, respectively) than container-grown seedlings with little variation between container volume treatments (Table `r pvetab3`). Overall, the *J*~max~ / *V*~cmax~ ratio was conserved across all treatments (1.6&pm;0.02), which is consistent with findings across many tree species [@medlyn1996optimal; @leuning1997scaling; @medlyn1999effects; @warren2003photosynthesis; @crous2008elevated]. Leaf dark respiration rates were not significantly different across soil volumes (Table `r pvetab3`). The *g*~1~ parameter, generated for each seedling from the Medlyn *et al.* [-@medlyn2011reconciling] optimal stomatal conductance model, was lowest in the free seedling treatment and was marginally different across soil volume treatments (Table `r pvetab3`).

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 MPa across all seedlings, respectively. Although *g*~s~ in free seedlings was generally higher than those in containers (Table `r pvetab3`, *P* < 0.001), the mean rates for all seedlings were high at 0.37 mol m^-2^ s^-1^ and did not change throughout the course of the experiment. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table `r pvetab1`). Combined, these indices provide strong evidence that water stress was not apparent on these seedlings throughout the experiment. Soil N at harvest was not different across soil volumes (mean of .045 %), with minimal decreases from pre-planting values (mean of = .049 %). This similarity indicates that nutrient leaching from free seedlings or from draining of containers following natural rainfall events did not differ between treatments.

##Whole-plant C balance
Modelled cumulative net leaf C gain over 120 days varied two-fold across containers sizes (25.6 - 55.1 g C). The free seedling control had five-fold greater modelled total C gain than the average of container treatments (213.6 - 42.7&pm;8.2 g C). For seedlings in containers, an average of 67 &pm;0.01 % of modelled C gain was not allocated to observed biomass, compared to 59.2 % for the free seedling control. Consequently, the CUE of containerized seedlings was lower than the free seedling control (Figure `r cuefig`). As a result, the observed reductions in leaf *A*~n~ with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the container size treatments.

@@

##Changes in growth and physiology under sink limitation
Soon after seedlings became established, both height and diameter growth were negatively affected by decreasing soil volume. Low growth rates led to large reductions in biomass in small containers, compared to freely rooted seedlings. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, close to the 43 % reported in the meta-analysis by Poorter *et al.* [-@poorter2012pot]. These growth reductions were expected, as the impedance of root growth can cause reductions in overall plant growth and activity [@mcconnaughay1991physical; @young1997mechanical]. It has been shown that roots subjected to environmental stress may send inhibitory signals to the shoots that affect *g*~s~, cell expansion, cell division and the rate of leaf appearance [@passioura2002soil]. Here, we observed a large divergence in leaf area between seedlings in containers and free seedlings through time, with the eventual cessation of new leaf growth in seedlings in small containers.

@@

\
With respect to nutrient availability, reductions in *A*~n~ were positively correlated with decreases in leaf N~f~ and leaf N~f~ was considerably reduced for seedlings in containers. Since reductions in leaf N~f~ were detected with TNC-free leaf mass, TNC dilution of leaf N was accounted for in all seedlings. It is possible that either physical root restriction or decreased supply reduced seedling N uptake in small containers. Although root N was on average higher in free seedlings at the end of the experiment, it was not consistently higher than the smallest container volume treatment. Unrestricted mycorrhizal recruitment could have instead facilitated the increases in leaf N~f~ in free seedlings, but this effect is unknown. Soil N also declined evenly across all treatments, providing no clear evidence for decreased supply between free seedlings and seedlings in containers. In these already low fertility soils, it is possible that seedlings in containers simply grew into increasing N limitation which negatively affected belowground sink strength. Although no clear feedback could be determined between the available soil N pool and decreases in leaf N~f~, the effects of belowground sink limitation on *A*~n~ of seedlings in containers was evident throughout the experiment.

\
As both rooting space and resources were finite in containers, the inability of seedlings to maintain the capacity of the belowground C sink resulted in the buildup of C assimilate in leaves. The feedback inhibition of *A*~n~ from starch accumulation has been proposed, yet it is still not known whether there is a starch threshold that triggers the down-regulation process [@nebauer2011photosynthesis]. Here, declines in *A*~max~ were correlated with higher starch content throughout the experiment. These results agree with a study on a deciduous conifer by Equiza *et al.* [-@equiza2006photosynthetic] where photosynthetic downregulation from reduced sink strength was correlated with starch content. As starch content in leaves of plants grown in the smallest containers was nearly double that of free seedlings in our study, this suggests the response of *A*~n~ to sink inhibition was regulated by this accumulation, as hypothesized. Overall, it is likely that both nutrient content and reduced sink strength played a role in observed photosynthetic down regulation, but future studies are needed to identify the specific mechanisms which prompted these feedbacks.

##Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to free seedlings, once ontogenetic drift was taken into account (Figure `r allocfig`a,b). A conserved ratio of fine root mass to leaf mass was observed across all treatments, suggesting that seedlings maintained homeostasis with biomass partitioning instead of a functional partitioning response to optimize limiting soil resources.

@@

\
**Table 2**. Responses of root characteristics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment. All values are from the final harvest. Values for fine root length density (FRLD) were only calculated for seedlings in containers as free seedlings had potentially unlimited soil volume to exploit. Different letters represent significant differences between treatments. The volume effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings, except for FRLD which represents only differences between seedlings in containers.
```{r table2, echo=FALSE, warning=FALSE}
dust(table2)%>%
  sprinkle_colnames("Volume (L)", 
                        "Root N (%)", 
                        "SRL (m g^-1^)",
                        "FRLD (m dm^-3^)") %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

@@

**Figure 5**. Photosynthetic capacity on a leaf mass basis (A~mass~), was correlated to both leaf starch and nitrogen content. (a) A~mass~ was negatively related to leaf starch (shown as percentage of dry mass), in particular for leaves with higher nitrogen content (N~mass~). (b) A~mass~ was positively related to N~mass~ (expressed as N per unit mass after subtracting TNC content). In both panels, each symbol is an individual measurement of A~mass~ (data across all campaigns are shown). Colors represent bins levels (n = 5) of both leaf starch and nitrogen grouped from low to high. Lines represents predictions, for each bin level, from the linear mixed effects model equation of A~mass~ as a function of starch and nitrogen. The marginal *R*^2^ (fixed effects only) was 0.37 and the conditional *R*^2^ (fixed and random effects) was 0.48 for the complete model. The equation for the full model is y = - 221.96 + 28.58(N) - 3.31(Starch) - 0.87(N*Starch).

>>>>

\
In the absence of a good understanding of the effects of sink limitation on physiological processes, most attempts to model sink limitation of growth to date have done so by abandoning the principle of mass balance [@fatichi2014moving]. This approach is unsatisfactory because mechanistic models should obey the laws of physics. Growth models must take into account several simultaneous processes affecting the C balance at any given time [@fourcaud2008plant], thus incomplete modelling approaches often lead to unaccounted-for C. Our goal in this paper was to investigate how the C balance of plants is modified by sink limitation, by experimentally imposing a sink limitation to growth.

@@

\
Forty-nine *Eucalyptus tereticornis* seedlings, 20 weeks old and approximately 40 cm tall in tube stock, were chosen from a single local Cumberland plain (western Sydney) cohort. Six additional seedlings were harvested before planting to measure initial leaf area and dry mass of leaves, stems and roots. Six container volumes were used ranging from 5 L to 35 L, with a 22.5 cm diameter, and lengths ranging from 15 to 100 cm. Containers were constructed of PVC pipe and were filled with local top soil. Soil in each container was packed to achieve a target soil bulk density that matched local soil conditions of 1.7 g cm^-3^. An Imidacloprid (BAYER CropScience) insecticide tablet was planted 5 cm below the roots of each seedling. Containers were planted flush with the soil surface inside metal sleeves, designed to minimize excess air space between the container and outside soil while also allowing for container removal. This allowed for soil temperatures in containers to reflect conditions of naturally planted (Free) seedlings. Each experimental block (n=7) contained a complete replicate set of six container volumes as well as one Free seedling, with 1 m^2^ spacing. For each Free seedling, used as the control, a 1 m^2^ subplot was excavated to the hard layer and replaced with the same soil used in each container. A border of root exclusion material was buried 0.25 m deep and extended 0.25 m above the ground surface around each subplot to exclude local vegetation, which was further kept out by periodic weeding.

@@

##Growth and morphology metrics
Seedlings were planted in summer (January 21^st^ 2013) and stem height, diameter at 15 cm and leaf count were measured weekly thereafter. After four months, a full biomass harvest was completed and the experiment ended (May 21^st^ 2013). Dry mass of leaves, stems, roots and total leaf area (LI-3100C Area Meter; LI-COR, Lincoln, NE, USA) were measured for each seedling. Mean individual leaf area for each harvested seedling was calculated by dividing total measured leaf area by total leaf count of only fully expanded leaves. Mean individual leaf area was then used to interpolate total seedling leaf area through time with weekly leaf counts. Root mass was collected by removing the roots system and passing soil from each container through a 1 mm sieve, washing, separating into fine and coarse roots (<2 mm and >2 mm diameter, respectively) and then drying to a constant mass. Roots of seedlings in containers were not considered pot bound, as matted roots along the soil-container interface were not observed. Roots from the Free seedlings were collected by excavating each 1 m^2^ subplot to the hard layer and keeping only roots within the subplot. For each seedling, a sub-sample of washed fine roots was analyzed for root length using WhinoRhizo software (Regent Instruments Inc., Quebec, QC, Canada). Specific root length (SRL) is reported as the root length divided by the dry mass of each sub-sample (m g^-1^). Fine root length density (FRLD) for seedlings in containers is reported as the total fine root length divided by the volume of each container (m dm^-3^).

@@

##Growth and morphology
Plant height, diameter and leaf area diverged between container volumes soon after start of the experiment (Figure `r allomfig`a-c). First to diverge was seedling leaf area, which significantly differed among soil volumes (*P* < 0.029) during the 5^th^ week of the experiment. Significant differences were then observed in both height (8^th^ week) and then diameter (9^th^ week) significantly among soil volumes (*P* < 0.045 & 0.035, respectively). The large reductions in height gain and total leaf area in smaller compared to larger containers continued throughout the experiment. In this field study, colder temperatures and reductions in total PPFD per day (Figure `r airvarsfig`) likely slowed the growth of seedlings in the final weeks of the experiment. Seedlings maintained diameter growth throughout the experiment, although growth was marginal in smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (*P* < 0.001). Increases in both final stem diameter and cumulative leaf area (both *P* < 0.001) were found with increasing soil volume. We did not find a continuous effect of container size; rather, plant size was similar among smaller containers, and differences were driven mainly by the largest container and the Free seedling treatments.

\
Total seedling biomass at harvest was significantly different across container volumes (*P* < 0.001) and between container treatments and Free seedlings (*P* < 0.001, Table `r pvetab1`). On average, harvested biomass of Free seedlings was 84% higher than that of seedlings in containers. Plant biomass was positively correlated with total leaf area across all treatments (*R*^2^ = 0.97, *P* < 0.001). Differences in biomass partitioning to leaves, stems and roots were not different across soil volumes after variation in seedling biomass across treatments was factored in the analysis (Figure `r allocfig`a,b). The final harvest root:shoot biomass ratio and  was conserved across all treatments, with a slightly higher shoot than root mass ($\overline{x}$ = 0.904, 95% CI = [0.846,1.119]). The ratio of leaf to fine root mass was also not different between treatments (Figure `r allocfig`c).

\
Overall, SRL was higher in seedlings in containers compared to Free seedlings but only in some of the container size treatments (Table `r pvetab2`, *P* = 0.009). Fine root length density was significantly higher in the two smallest container sizes and was the lowest in the largest container size (Table `r pvetab2`, *P* < 0.001). Over the duration of the experiment SLA~f~ was higher in Free seedlings, but was not different across container sizes (Table `r pvetab1`, *P* < 0.001) and this pattern was evident beginning in the first gas exchange measurement campaign (*P* < 0.001).

##Leaf and root chemistry
Leaf N~f~ was significantly higher in Free seedlings and the largest container volume compared to the smaller container volumes at the onset of gas exchange measurements (6th week, *P* < 0.001). Throughout the remainder of the experiment the smallest container volume had significantly lower leaf N~f~ than other soil volumes, while Free seedlings maintained the highest leaf N~f~ (Table `r pvetab1`, *P* < 0.001). Leaf starch content in the smallest container was ca. double  that of Free seedlings (*P* = 0.039), while leaf soluble sugars did not differ across treatments throughout the experiment (Table `r pvetab1`). Differences in leaf starch between the Free seedling and the smallest container were evident during the first gas exchange campaign (*P* = 0.001). Root N was higher in Free seedlings compared to seedlings in containers but only for some of the container size treatments (Table `r pvetab2`).

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the Free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ (Figure `r asatfig`) and *A*~max~ (Table `r pvetab3`) were consistently higher in Free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf starch and leaf N~f~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis was highly correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~max~ to either leaf N~f~ or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~max~ was higher when leaf N~f~ was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~max~  was also lower when leaf starch was high as higher leaf N~f~ often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, *A*~max~ was positively correlated with final harvest biomass across all seedlings (*P* < 0.001).

\
Both *J*~max~ and *V*~cmax~ were significantly higher in Free seedlings (30 % and 26 %, respectively) than container-grown seedlings with little variation between container volume treatments (Table `r pvetab3`). Overall, the *J*~max~ / *V*~cmax~ ratio was conserved across all treatments (1.6&pm;0.02), which is consistent with findings across many tree species [@medlyn1996optimal; @leuning1997scaling; @medlyn1999effects; @warren2003photosynthesis; @crous2008elevated]. Leaf dark respiration rates were not significantly different across soil volumes (Table `r pvetab3`). The *g*~1~ parameter, generated for each seedling from the Medlyn *et al.* [-@medlyn2011reconciling] optimal stomatal conductance model, was lowest in the Free seedling treatment and was marginally different across soil volume treatments (Table `r pvetab3`).

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 MPa across all seedlings, respectively. Although *g*~s~ in Free seedlings was generally higher than those in containers (Table `r pvetab3`, *P* < 0.001), the mean rates for all seedlings were high at 0.37 mol m^-2^ s^-1^ and did not change throughout the course of the experiment. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table `r pvetab1`). Combined, these indices provide strong evidence that water stress was not apparent on these seedlings throughout the experiment. Soil N at harvest was not different across soil volumes (mean of .045 %), with minimal decreases from pre-planting values (mean of = .049 %). This similarity indicates that nutrient leaching from Free seedlings or from draining of containers following natural rainfall events did not differ between treatments.

##Whole-plant C balance
Modelled cumulative net leaf C gain over 120 days varied two-fold across containers sizes (25.6 - 55.1 g C). The Free seedling control had five-fold greater modelled total C gain than the average of container treatments (213.6 - 42.7&pm;8.2 g C). For seedlings in containers, an average of 67 &pm;0.01 % of modelled C gain was not allocated to observed biomass, compared to 59.2 % for the Free seedling control. Consequently, the CUE of containerized seedlings was lower than the Free seedling control (Figure `r cuefig`). As a result, the observed reductions in leaf *A*~n~ with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the container size treatments.

@@

##Changes in growth and physiology under sink limitation
Soon after seedlings became established, both height and diameter growth were negatively affected by decreasing soil volume. Low growth rates led to large reductions in biomass in small containers, compared to freely rooted seedlings. We analyzed the relationship between biomass growth and soil volume and found an increase of 34 % with a doubling of container volume, close to the 43 % reported in the meta-analysis by Poorter *et al.* [-@poorter2012pot]. These growth reductions were expected, as the impedance of root growth can cause reductions in overall plant growth and activity [@mcconnaughay1991physical; @young1997mechanical]. It has been shown that roots subjected to environmental stress may send inhibitory signals to the shoots that affect *g*~s~, cell expansion, cell division and the rate of leaf appearance [@passioura2002soil]. Here, we observed a large divergence in leaf area between seedlings in containers and Free seedlings through time, with the eventual cessation of new leaf growth in seedlings in small containers.

@@

\
With respect to nutrient availability, reductions in *A*~n~ were positively correlated with decreases in leaf N~f~ and leaf N~f~ was considerably reduced for seedlings in containers. Since reductions in leaf N~f~ were detected with TNC-free leaf mass, TNC dilution of leaf N was accounted for in all seedlings. It is possible that either physical root restriction or decreased supply reduced seedling N uptake in small containers. Although root N was on average higher in Free seedlings at the end of the experiment, it was not consistently higher than the smallest container volume treatment. Unrestricted mycorrhizal recruitment could have instead facilitated the increases in leaf N~f~ in Free seedlings, but this effect is unknown. Soil N also declined evenly across all treatments, providing no clear evidence for decreased supply between Free seedlings and seedlings in containers. In these already low fertility soils, it is possible that seedlings in containers simply grew into increasing N limitation which negatively affected belowground sink strength. Although no clear feedback could be determined between the available soil N pool and decreases in leaf N~f~, the effects of belowground sink limitation on *A*~n~ of seedlings in containers was evident throughout the experiment.

\
As both rooting space and resources were finite in containers, the inability of seedlings to maintain the capacity of the belowground C sink resulted in the buildup of C assimilate in leaves. The feedback inhibition of *A*~n~ from starch accumulation has been proposed, yet it is still not known whether there is a starch threshold that triggers the down-regulation process [@nebauer2011photosynthesis]. Here, declines in *A*~max~ were correlated with higher starch content throughout the experiment. These results agree with a study on a deciduous conifer by Equiza *et al.* [-@equiza2006photosynthetic] where photosynthetic downregulation from reduced sink strength was correlated with starch content. As starch content in leaves of plants grown in the smallest containers was nearly double that of Free seedlings in our study, this suggests the response of *A*~n~ to sink inhibition was regulated by this accumulation, as hypothesized. Overall, it is likely that both nutrient content and reduced sink strength played a role in observed photosynthetic down regulation, but future studies are needed to identify the specific mechanisms which prompted these feedbacks.

##Biomass partitioning under sink limitation
As biomass partitioning is likely controlled by the source and sink strength of all organs [@poorter2012biomass], it was important to determine which tissue components were most affected by the container size treatments. It was necessary to distinguish if growth was affected beyond ontogenetic constraints, by correcting for size, as biomass distribution is strongly size-dependent [@gould1966allometry; @lleonart2000removing]. In this study, there was no significant difference in root, leaf, or stem biomass partitioning with reduced soil volume compared to Free seedlings, once ontogenetic drift was taken into account (Figure `r allocfig`a,b). A conserved ratio of fine root mass to leaf mass was observed across all treatments, suggesting that seedlings maintained homeostasis with biomass partitioning instead of a functional partitioning response to optimize limiting soil resources.

@@

\
**Table 2**. Responses of root characteristics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment. All values are from the final harvest. Values for fine root length density (FRLD) were only calculated for seedlings in containers as Free seedlings had potentially unlimited soil volume to exploit. Different letters represent significant differences between treatments. The volume effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings, except for FRLD which represents only differences between seedlings in containers.
```{r table2, echo=FALSE, warning=FALSE}
dust(table2)%>%
  sprinkle_colnames("Volume (L)", 
                        "Root N (%)", 
                        "SRL (m g^-1^)",
                        "FRLD (m dm^-3^)") %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

@@

**Figure 5**. Photosynthetic capacity on a leaf mass basis (A~mass~, measured at saturating light and [CO~2~]) was correlated to both leaf starch and nitrogen content. (a) A~mass~ was negatively related to leaf starch (shown as percentage of dry mass), in particular for leaves with higher nitrogen content (N~mass~). (b) A~mass~ was positively related to N~mass~ (expressed as N per unit mass after subtracting TNC content). In both panels, each symbol is an individual measurement of A~mass~ (data across all campaigns are shown). Colors represent bins levels (n = 5) of both leaf starch and nitrogen grouped from low to high. Lines represents predictions, for each bin level, from the linear mixed effects model equation of A~mass~ as a function of starch and nitrogen. The marginal *R*^2^ (fixed effects only) was 0.37 and the conditional *R*^2^ (fixed and random effects) was 0.48 for the complete model. The equation for the full model is y = - 221.96 + 28.58(N) - 3.31(Starch) - 0.87(N*Starch).
